---
title: "hw6"
output: html_document
date: "2024-12-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(modelr)
library(broom)
library(crossval)
library(mgcv)
```

---

### Problem 2

```{r}
#load the data
homicides <- read_csv("./data/homicide-data.csv")
```

```{r}
homicides <- homicides %>%
  mutate(
    city_state = str_c(city, ", ", state),
    solved_binary = if_else(disposition == "Closed by arrest", 1, 0),
    victim_age = as.numeric(victim_age)
  ) %>%
  filter(
    !city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"),
    victim_race %in% c("White", "Black"),
    !is.na(victim_age)
  )
head(homicides)
```

```{r}
#only baltimore data
baltimore <- homicides %>% 
  filter(city_state == "Baltimore, MD")

#ffit logistic regression model
baltimore_glm <- glm(
  solved_binary ~ victim_age + victim_sex + victim_race,
  family = binomial(),
  data = baltimore
)

#obtain odds ratio for male victims
baltimore_or <- broom::tidy(baltimore_glm) %>%
  filter(term == "victim_sexMale") %>%
  mutate(
    or = exp(estimate),
    ci_lower = exp(estimate - 1.96 * std.error),
    ci_upper = exp(estimate + 1.96 * std.error)
  )
baltimore_or 
```

```{r}
#logistic regression for all cities
cities <- homicides %>%
  group_by(city_state) %>%
  nest() %>%
  mutate(
    model = map(data, ~ glm(
      solved_binary ~ victim_age + victim_sex + victim_race,
      family = binomial(),
      data = .
    )),
    tidy_model = purrr::map(model, broom::tidy)
  ) %>%
  unnest(tidy_model) %>%
  filter(term == "victim_sexMale") %>%
  mutate(
    or = exp(estimate),
    ci_lower = exp(estimate - 1.96 * std.error),
    ci_upper = exp(estimate + 1.96 * std.error)
  ) %>%
  select(city_state, or, ci_lower, ci_upper)

cities %>%
  slice(1:5) %>%
  knitr::kable(digits = 3)
```

```{r}
#plot
cities %>%
  ggplot(aes(x = fct_reorder(city_state, or), y = or)) +
  geom_point() +
  geom_errorbar(aes(ymin = ci_lower, ymax = ci_upper), width = 0.2) +
  coord_flip() +
  labs(
    title = "Adjusted Odds Ratios for Solving Homicides",
    x = "City, State",
    y = "Adjusted Odds Ratio"
  ) +
  theme_minimal()

```
Most cities smaller than OR of 1, indicating that after controlling for victim age and race, the likelihood of a crime involving a male victim being resolved is lower than that of a crime involving a female victim.

---

### Problem 3

```{r}
birthweight <- read_csv("./data/birthweight.csv")
```
```{r}
birthweight <- birthweight %>%
  janitor::clean_names() %>%
  mutate(
    babysex = factor(babysex, levels = c(1,2), labels = c("male", "female")),
    malform = case_when(malform == 0 ~ "absent", malform == 1 ~ "present"),
    mrace = factor(mrace, levels = c(1,2,3,4,8), labels = c("White", "Black", "Asian", "Puerto Rican", "Other")),
    frace = factor(frace, levels = c(1,2,3,4,8,9), labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknow"))) %>%
  drop_na() 
```

```{r}
#average birthweight by baby's sex
birthweight %>%
  group_by(babysex) %>%
  summarise(
    avg_birthweight = mean(bwt, na.rm = TRUE),
    count = n()
  )
```


My model trying to capture the most critical factors influencing birthweight while avoiding overfitting and ensuring interpretability. To achieve this goal I believe following predictors are important, Gestational Age, Maternal Weight Gain, Baby’s Sex, and Malformations. The reason are following:
Gestational age is universally recognized as a primary determinant of birthweight.
Maternal Weight Gain is the weight a mother gains during pregnancy, which I beleive is directly linked to the baby’s growth and weight.
Baby’s Sex: Male babies tend to have slightly higher birthweights on average compared to female babies, this being verified by the code above. Male has 3157.583 avg birthweight while female has 3068.812 avg birthweight.
Malformations: Congenital malformations may adversely affect fetal growth, leading to lower birthweights.

```{r}
#fit the proposed regression model
mymodel <- lm(bwt ~ gaweeks + wtgain + babysex + malform, data = birthweight)


#add some predictions and residuals
birthweight <- birthweight %>%
  modelr::add_predictions(mymodel, var = "fitted") %>%
  modelr::add_residuals(mymodel, var = "residuals")

```

```{r}
#plot
birthweight %>%
  ggplot(aes(x = fitted, y = residuals)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "loess", color = "blue", se = FALSE) +
  labs(
    title = "Residuals vs Fitted Values",
    x = "Fitted Values",
    y = "Residuals"
  ) +
  theme_minimal()
```

```{r}
#crossvalidation
cv_data <- birthweight %>%
  crossv_mc(100) %>%
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
```

```{r}
cv_results <- cv_data %>%
  mutate(
    mymodel = map(train, ~ lm(bwt ~ gaweeks + wtgain + babysex + malform, data = .)),
    model1 = map(train, ~ lm(bwt ~ blength + gaweeks, data = .)),
    model2 = map(train, ~ lm(bwt ~ bhead * blength * babysex, data = .)),
    rmsemy = map2_dbl(mymodel, test, ~ rmse(model = .x, data = .y)),
    rmse1 = map2_dbl(model1, test, ~ rmse(model = .x, data = .y)),
    rmse2 = map2_dbl(model2, test, ~ rmse(model = .x, data = .y))
  )
```

```{r}
#summarize the data
cv_summary <- cv_results %>%
  summarise(
    mean_rmsemy = mean(rmsemy),
    mean_rmse1 = mean(rmse1),
    mean_rmse2 = mean(rmse2)
  )
```

```{r}
cv_results %>%
  select(starts_with("rmse")) %>%
  pivot_longer(
    cols = everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>%
  ggplot(aes(x = model, y = rmse)) +
  geom_violin(fill = "lightblue") +
  labs(
    title = "RMSE Comparison of Models",
    x = "Models",
    y = "RMSE"
  ) 
```

